---
title: "sim_learn"
author: "AJB"
date: "2023-10-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Github

```{r}

library(usethis)
usethis::use_github()
```



# Causal Diagram

```{r}
#install.packages("dagitty")
library(dagitty)

```


```{r}

simdi <- dagitty("dag {
                Y <- W -> X
                X -> Y
                e -> X
                }")

# Manually set the coordinates for each node
# The coordinates function expects a named list of x,y pairs

coordinates(simdi) <- list(
  x = c(X = 0, Y = 1, e = -1, W = 1),
  y = c(X = 0, Y = 0, e = 0, W = -1)
)
# Now plot the DAG with the specified coordinates
# Overlay bold labels

plot(simdi, textsize = 5)


```

# Creating Data

## installing tidyverse

```{r}
#install.packages("tidyverse")

```

## Creating data that follows Causal Diagram

```{r}

library(tidyverse)

# W is continuous; X is binary; Y is continuous; e will be random noise

# If we want the random data to be the same every time, set a seed
set.seed(1000)

# tibble() creates a tibble; it's like a data.frame
# We must pick a number of observations, here 200

d <- tibble(W = runif(200, 0, .1)) %>% # only go from 0 to 0.1
  mutate(X = runif(200) < .2 + W) %>% # runif(200) is what creates e
  mutate(Y = 3*X + W + rnorm(200)) # True effect of X on Y is 3



```

We can go one step further. Not only do we want to randomly create this data, we are going to want to do so a bunch of times as we repeatedly run our simulation. Any time we have code we want to run a bunch of times, it’s often a good idea to put it in its own function, which we can repeatedly call. We’ll also give this function an argument that lets us pick different sample sizes.

```{r}
# library(tidyverse)
# Make sure the seed goes OUTSIDE the function. It makes the random 
# data the same every time, but we want DIFFERENT results each time 
# we run it (but the same set of different results, thus the seed)
set.seed(1000)

# Make a function with the function() function. The "N = 200" argument 
# gives it an argument N that we'll use for sample size. The "=200" sets
# the default sample size to 200

create_data <- function(N = 200){
  d <- tibble(W = runif(N, 0, 0.1)) %>%
    mutate(X = runif(N) < 0.2 + W) %>%
    mutate(Y = 3*X + W + rnorm(N))
  
  # Use return() to send our created data back
  return(d)
}

# Run our function!

d550 <- create_data(550)



```


# Creating Panel Data

```{r}
# library(tidyverse)
set.seed(1000)

# N for number of individuals, T for time periods

create_panel_data <- function(N = 200, T = 10){
  
  # Create individual IDs with the crossing()
  # function, which makes every combination of two vectors
  # (if you want some to be incomplete, drop them later)
  
  panel_data <- crossing(ID = 1:N, t = 1:T) %>%
    
    # And individual/time-varying data
    
    # (n() means "the number of rows in this data"):
    
    mutate(W1 = runif(n(), 0, 0.1))
  
  
    # Now an individual-specific characteristic
  
    indiv_data <- tibble(ID = 1:N, W2 = rnorm(N))
    
    # Join them
    
    panel_data <- panel_data %>%
      full_join(indiv_data, by = 'ID') %>%
      
      # Create X, caused by W1 and W2
      
      mutate(X = 2*W1 + 1.5*W2 + rnorm(n())) %>%
      
      # And create Y. The true effect of X on Y is 3
      
      # But W1 and W2 have causal effects too
      
      mutate(Y = 3*X + W1 - 2*W2 + rnorm(n()))
    
    
    return(panel_data)
    
    
}
```


```{r}
# take it for a spin

d_pan <- create_panel_data(100, 5)


```


# Creating Heteroskedasticity

```{r}

# libray(tidyverse)

set.seed(1000)

create_het_data <- function(N = 200){
  
  d <- tibble(X = runif(N)) %>%
    
    # Let the standard deviation of the error
    # be related to X. Heteroskedasticity!
    
    mutate(Y = 3*X + rnorm(N, sd = X*5))
  
  
    return(d)
}



```

```{r}
# runnin it

d_het <- create_het_data(500)



```

```{r}
# visualizing a class "fan" shape we've created

plot(d_het)



```






